---
title: "HW3"
author: "林茂廷"
date: "5/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
.answer {
  border-style: dashed;
}
```


## 1 Equivalence between demean and first-difference
考慮如下的固定效果模型：
$$y_{it}=x_{it}'\beta+c_{i}+\epsilon_{it}.$$
使用demean估計時，我們是以下式執行OLS估計： $$y_{it}-\bar{y}_{i.}=(x_{it}-\bar{x}_{i.})'\beta+(\epsilon_{it}-\bar{\epsilon}_{i.}).$$ 
另一種方式是使用一階差分，也就是對下式執行OLS估計：
$$\Delta y_{it}=\Delta x_{it}'\beta+\Delta \epsilon_{it}.$$
若資料只有t=0,1兩期，說明使用demean估計與一階差分估計會得到相同結果（可以假設x為單一變數讓證明更簡單）。


<div class="answer">
Based on the demean approach, 
$$
\hat{\beta}=
\frac
{\sum_{it}[(x_{it}-\bar{x}_{i.})(y_{it}-\bar{y}_{i.})]}
{\sum_{it}(x_{it}-\bar{x}_{i.})^2}
$$
Where data only last two periods, say 0 and 1, for $w\in (x,y)$
$$
\begin{array}{lcl}
w_{it}-\bar{w}_{i.}&=&w_{it}-\frac{w_{i0}+w_{i1}}{2}\\
&=&
\begin{cases}
\frac{1}{2}(w_{i0}-w_{i1})\mbox{ if } i=0\\
\frac{1}{2}(w_{i1}-w_{i0})\mbox{ if } i=1
\end{cases}
\end{array}
$$
Hence
$$
\begin{array}{lcl}
\sum_{it}[(x_{it}-\bar{x}_{i.})(y_{it}-\bar{y}_{i.})]&=&\sum_{i,t=0}[(x_{it}-\bar{x}_{i.})(y_{it}-\bar{y}_{i.})]+\\
& & \sum_{i,t=1}[(x_{it}-\bar{x}_{i.})(y_{it}-\bar{y}_{i.})]\\
&=& \sum_i [\frac{1}{4}(x_{i0}-x_{i1})(y_{i0}-y_{i1})]+\\
& & \sum_i [\frac{1}{4}(x_{i1}-x_{i0})(y_{i1}-y_{i0})]\\
&=& \sum_i [\frac{1}{2}(x_{i1}-x_{i0})(y_{i1}-y_{i0})]
\end{array}
$$
Similary,
$$
\sum_{it}(x_{it}-\bar{x}_{i.})^2=\sum_i [\frac{1}{2}(x_{i1}-x_{i0})^2].
$$
Together,
$$
\hat{\beta}=\frac{\sum_i [(x_{i1}-x_{i0})(y_{i1}-y_{i0})]}{\sum_i [(x_{i1}-x_{i0})^2]},
$$
which is exactly the same as the OLS estimator under the first-difference approach.
</div>


## 2 廣義固定效果
考慮以下廣義固定效果模型： 
$$y_{it}=x_{it}'\beta+c_{i}+\delta_{t}+\epsilon_{it},$$
定義以下平均表示符號：
$$
\begin{array}{lcl}
\bar{w}_{i.}&=&\frac{\sum_{t=1}^T w_{it}}{T}\\
\bar{w}_{.t}&=&\frac{\sum_{i=1}^N w_{it}}{N}\\
\bar{\bar{w}}&=&\frac{\sum_i\sum_t w_{it}}{NT}
\end{array}
$$
考慮如下的demean function 
$$
D(w_{it}|\bar{w}_{i.},\bar{w}_{.t},\bar{\bar{w}})=w_{it}-\bar{w}_{i.}-\bar{w}_{.t}+\bar{\bar{w}}
$$
說明對廣義固定效果模型左右式一起進行$D(.)$函數的demean轉換可以去除$c_i$及$\delta_t$兩個固定效果。


<div class="answer">
$$
\begin{array}{lcl}
D(y_{it}|\bar{y}_{i.},\bar{y}_{.t},\bar{\bar{y}})&=&y_{it}-\bar{y}_{i.}-\bar{y}_{.t}+\bar{\bar{y}}\\
&=& x_{it}'\beta+c_i+\delta_t+\epsilon_{it}\\
& & -(\bar{x}_{i.}'\beta+c_i+\sum_t\delta_t/T+\sum_t\epsilon_{it}/T)\\
& & -(\bar{x}_{.t}'\beta+\sum_ic_i/N+\delta_t+\sum_i\epsilon_{it}/N)\\
& & +(\bar{\bar{x}}'\beta+\sum_ic_i/N+\sum_t\delta_t/T+\sum_i\sum_t\epsilon_{it}/NT)\\
&=& D(x_{it}|\bar{x}_{i.},\bar{x}_{.t},\bar{\bar{x}})'\beta+D(\epsilon_{it}|\bar{\epsilon}_{i.},\bar{\epsilon}_{.t},\bar{\bar{\epsilon}})
\end{array}
$$
</div>
 

## 3 Panel data的迴歸殘差
考慮迴歸模型：
$$
y_{it}=\beta'x_{it}+\epsilon_{it}
$$
其中i代表某一家戶單位，t代表資料年份。

研究者將$\epsilon_{it}$訊息拆解成

  * $\alpha_i=\frac{\sum_{t=1}^T \epsilon_{it}}{T}$，及
  
  * $v_{it}=\epsilon_{it}-\alpha_i$為拆解後殘差。
  
（1）證明$\frac{\sum_{it} v_{it}\alpha_{i}}{NT}=0$。（註：當兩隨機變數w及z，其平均值為0時，$E(wz)=0$表示兩者covariance為0，即兩者無線性關連，故此題說明研究者的拆解方式會保證兩者sample covariance為0，有樣本上的零線性關連）

<div class="answer">
$$
\begin{array}{lcl}
\frac{\sum_{it} v_{it}\alpha_{i}}{NT}&=&
\frac{\sum_{it} (\epsilon_{it}-\alpha_{i})\alpha_{i}}{NT}\\
&=& \frac{\sum_i\alpha_{i}(\sum_t(\epsilon_{it}-\alpha_{i}))}{NT}
\end{array},
$$
where
$$
\sum_t(\epsilon_{it}-\alpha_{i})=T\alpha_i-T\alpha_i=0.
$$
Therefore, $\frac{\sum_{it} v_{it}\alpha_{i}}{NT}=0$.
</div>

（2）說明$E(v_{it}\alpha_i)=0$。（說理過程所使用的假設及定理都必需清楚寫下） 

<div class="answer">
According to Law of Large Number, 
$$
\sum_{it} v_{it}\alpha_{i}/NT \stackrel{P}{\rightarrow} E(v_{it}\alpha_{i}).
$$
Since $\sum_{it} v_{it}\alpha_{i}/T=0$ all the time, $E(v_{it}\alpha_i)=0$.
</div>

令$\sigma_{\alpha}^2=V(\alpha_i)$，$\sigma_{v}^2=V(v_{it})$，及$\sigma_{\alpha,v}=cov(\alpha_i,v_{it})$

（3）說明$cov(\alpha_i,v_{it})=0$且推導$V(\epsilon_{it})$，及$cov(\epsilon_{it},\epsilon_{is})$當$t\neq s$。

<div class="answer">
$cov(\alpha_i,v_{it})$:
$$
\begin{array}{lcl}
cov(\alpha_i,v_{it})&=&E(v_{it}\alpha_i)-E(v_{it})E(\alpha_i)\\
&=&0,
\end{array}
$$
due to the fact that $E(v_{it}\alpha_i)=0$ and $E(\alpha_i)=\sum_t E(\epsilon_{it})/T=0$ (which results from $E(\epsilon_{it})=0$).

$V(\epsilon_{it})$:
$$\begin{array}{lcl}
V(\epsilon_{it}) &=& cov(\alpha_i+v_{it},\alpha_i+v_{it})\\
&=& V(\alpha_i)+V(v_{it})+2cov(\alpha_i,v_{it})\\
& & \mbox{Because }cov(\alpha_i,v_{it})=0,\\
&=& V(\alpha_i)+V(v_{it})=\sigma_{\alpha}^2+\sigma_v^2 
\end{array}$$
$cov(\epsilon_{it},\epsilon_{is})$:
$$\begin{array}{lcl}
cov(\epsilon_{it},\epsilon_{is})&=&cov(\alpha_i+v_{it},\alpha_i+v_{is})\\
&=& V(\alpha_i)+cov(v_{it},v_{is})+cov(\alpha_i,v_{it})+cov(\alpha_i,v_{is}),\\
& & \mbox{Becasue }E(v_{it}\alpha_i)=0,\\ 
& & cov(\alpha_i,v_{it})=cov(\alpha_i,v_{is})=0\\
&=& V(\alpha_i)+cov(v_{it},v_{is})\\
&=& \sigma_{\alpha}^2+cov(v_{it},v_{is}).
\end{array}$$
As to the $cov(v_{it},v_{is})$, it requires more assumptions to deduct further. In random effect model, it is simply assumed that $cov(v_{it},v_{is})=0$ when $t\neq s$.
</div>

由於家戶單位為隨機抽出，我們可以假設對任意的s及t，$cov(\epsilon_{it},\epsilon_{js})=0$當$i\neq j$（即任兩不同家戶的迴歸模型殘差無關）。

（4）令
$$
\begin{array}{lcl}
\mathbf{\epsilon} &=&
\begin{pmatrix}
\mathbf{\epsilon}_1 \\ \mathbf{\epsilon}_2 \\ \vdots \\ \mathbf{\epsilon}_N
\end{pmatrix}
\mbox{ 其中 }
\mathbf{\epsilon}_{i}=
\begin{pmatrix}
\epsilon_{i1}\\ \epsilon_{i2}\\ \vdots \\ \epsilon_{iT}
\end{pmatrix}
\mbox{ 為第i個家戶單位的殘差向量}\\
\Omega&=&E(\epsilon\epsilon^T)
\end{array}
$$
請推導$\Omega$矩陣內容（此即為Random effect模型的殘差共變異矩陣）。推導步驟若有用到任何假設，必需清楚標明。

<div class="answer">
Due to random sampling across households, we only need to focus on $E(\epsilon_j\epsilon_j^T)$, that is the covariance matrix from the j-th sampled household. 

* $E(\epsilon_j\epsilon_j^T)$:

$$\begin{array}{lcl}
E(\epsilon_j\epsilon_j^T)&=&
E\begin{bmatrix}
\epsilon_{j1}^2 & \epsilon_{j1}\epsilon_{j2} & \dots & \epsilon_{j1}\epsilon_{jT}\\
\epsilon_{j2}\epsilon_{j1} & \epsilon_{j2}^2 & \dots & \epsilon_{j2}\epsilon_{jT}\\
\vdots & & \ddots & \\
\epsilon_{jT}\epsilon_{j1} & & & \epsilon_{jT}^2
\end{bmatrix}\\
&& \mbox{Base on the answer from the previous question, }\\
&=&\begin{bmatrix}
\sigma_{\alpha}^2+\sigma_v^2 & \sigma_{\alpha}^2 & \dots & \sigma_{\alpha}^2\\
\sigma_{\alpha}^2 & \sigma_{\alpha}^2+\sigma_v^2 & \dots & \sigma_{\alpha}^2\\
\vdots & & \ddots & \\
\sigma_{\alpha}^2 & & & \sigma_{\alpha}^2+\sigma_v^2
\end{bmatrix}\stackrel{denoted}{=}\Omega_d.
\end{array}$$
Then
$$
\Omega=\begin{bmatrix}
\Omega_d & 0 &\dots & 0\\
0 & \Omega_d & \dots & 0\\
\vdots & & \ddots & \\
0 & & & \Omega_d
\end{bmatrix}.
$$

</div>
（5）請問Fixed effect model的殘差指的是什麼？它的共變異矩陣在前述設定下會長怎麼樣？


<div class="answer">
For fixed effect model, $\alpha_i$ is eventually removed. Therefore, it is the covariance matrix of $\mathbf{v}$ that matters, where $\mathbf{v}$ is the stacking of all  $\{v_{it}\}$.

Due to random sampling across households, we only need to focus on $E(v_jv_j^T)$, that is the covariance matrix from the j-th sampled household. 

* $E(v_jv_j^T)$:
$$\begin{array}{lcl}
E(v_jv_j^T)&=&
E\begin{bmatrix}
v_{j1}^2 & v_{j1}v_{j2} & \dots & v_{j1}v_{jT}\\
v_{j2}v_{j1} & v_{j2}^2 & \dots & v_{j2}v_{jT}\\
\vdots & & \ddots & \\
v_{jT}v_{j1} & & & v_{jT}^2
\end{bmatrix}\\
& &\mbox{Assume cov(v_{jt},v_{js})=0 as in RE model and denote $v(v_{jt})=\sigma_v^2$}\\
&=& \begin{bmatrix}
\sigma_v^2 & 0 & \dots & 0\\
0 & \sigma_v^2 & \dots & 0\\
\vdots & & \ddots & \\
0 & & & \sigma_v^2
\end{bmatrix}\stackrel{denoted}{=}\Omega_v.\\
\end{array}$$
Then
$$
\begin{array}{lcl}
E(\mathbf{v}\mathbf{v}^T)=
\begin{bmatrix}
\Omega_v & 0 &\dots & 0\\
0 & \Omega_v & \dots & 0\\
\vdots & & \ddots & \\
0 & & & \Omega_v
\end{bmatrix}
\end{array}
$$
</div>


## 4 DiD
考慮以下新聞：2015年8月北市3項老人福利政策施予對象以年滿65歲以上老人為主，只要持有敬老悠遊卡就可享有乘車優惠。

我們想了解新政策是否增進老人行動力，$y$變數為某種行動力指標，$x$為代表「持有敬老悠遊卡就可享有乘車優惠」的效應解釋變數（若受此政策影響則為1，反之為0）。

（1）假設資料均來自於2016年北市與新北市老年人抽樣，考慮如下迴歸模型：
$$
y_i=\beta_0+\beta^*x_i+\epsilon_i
$$
說明為何$\beta^*$的OLS估計不會正確。

<div class="answer">
In this case, 
$$
\beta^*=E(y_i|x_i=1)-E(y_i|x_i=0), 
$$
which is basically the elderly mobility comparision between Taipei and New Taipei cities in 2016. However, even without the welfare policy, these two cities are different. Disparities can be in their average ages, family structures, income levels, etc. The initial condition difference contaminates the estimated $\beta^*$ result.
</div>

假設資料來自於2014及2016年，分別以`t=0`及`t=1`表示。另外，以`s=0`表示新北市，`s=1`表示台北市。考慮如下迴歸模型：
$$
y_{st,i}=\beta_0+\beta^*x_{st,i}+\epsilon_{st,i}
$$
（2）說明$\beta^*$的OLS估計式是在計算「2016年台北市老人平均活動力」減去「2014年台北市，及2014/2016兩年新北市老人平均活動力」
。

<div class="answer">
In this case, $x_{st,i}=1$ represents Taipei city in 2016, while $x_{st,i}=0$ represents other situations which include New Taipei City in both 2014 and 2016 as well as Taipei City in 2014. Therefore, 
$$
\beta^*=E(y_i|x_i=1)-E(y_i|x_i=0), 
$$
which is basically the elderly mobility comparision between those two groups. 
</div>

（3）承接上題。說明OLS估計不能代表「悠遊卡福利政策」影響。

<div class="answer">
As one group contains New Taipei city elderly, but not the other group, the initial condition bias that we worry in question (1) still exists. 
</div>

接著考慮如下迴歸模型：
$$
y_{st,i}=\beta_0+\alpha_1D1_s+\beta^*x_{st,i}+\epsilon_{st,i}
$$
其中$D1$為代表「台北市」的虛擬變數。

（4）說明$\beta^*$的OLS估計式是在計算「2016年台北市老人平均活動力」減去「2014年台北市老人平均活動力」。
。

<div class="answer">
Taipei City in 2016 can be expressed as $D1=1,X=1$, while Taipei City in 2014 as $D1=1,X=0$. The mobility comparision is, therefore,
$$
E(y|D1=1,X=1)-E(y|D1=1,X=0)=\beta^*
$$
</div>

（5）承接上題。說明OLS估計不能代表「悠遊卡福利政策」影響。

<div class="answer">
Even within Taipei city, without the welfare policy, the elderly mobility can still change between 2014 and 2016. The average age can get higher; the income level can get higher; and so on so forth. Therefore, the comparison in the previous question does not necessary represent the effect of the welfare policy.
</div>

接著考慮如下迴歸模型：
$$
y_{st,i}=\beta_0+\delta_1B1_t+\beta^*x_{st,i}+\epsilon_{st,i}
$$
其中$B1$為代表「2016年」的虛擬變數。


（6）說明$\beta^*$的OLS估計式是在計算「2016年台北市老人平均活動力」減去「2016年新北市老人平均活動力」。
。

<div class="answer">
Taipei City in 2016 can be expressed as $B1=1,X=1$, while New Taipei City in 2016 as $B1=1,X=0$. The mobility comparision is, therefore,
$$
E(y|B1=1,X=1)-E(y|B1=1,X=0)=\beta^*
$$
</div>

（7）承接上題。說明OLS估計不能代表「悠遊卡福利政策」影響。

<div class="answer">
This estimate still compares Taipei city and New Taipei city elderly directly without considering the fact that they are demographically different at the beginning. 
</div>

前面的題目點出政策效果的研究難處在於處理「比較對象立足點差異」及「政策前後的時間差異可能代表其他非政策的影響」，為解決這問題考慮以下的迴歸模型（即DiD）：

$$
y_{st,i}=\beta_0+\alpha_1 D1_s+\delta_1 B1_t+\beta^*x_{st,i}+\epsilon_{st,i}
$$

（8）說明第6小題「2016年台北市老人平均活動力」減去「2016年新北市老人平均活動力」不只帶有政策效果$\beta^*$，還帶有兩市立足點（即沒有政策效果下）的差異$\alpha_1$。

<div class="answer">
In DiD model, the comparison between these two groups is
$$\begin{array}{lcl}
E(y|D1=1,B1=1,X=1)-E(y|D1=0,B1=1,X=0)&=&(\beta_0+\alpha_1+\delta_1+\beta^*)\\
& & -(\beta_0+\delta_1)\\
&=& \alpha_1+\beta^*
\end{array}$$
</div>

（9）說明$\beta^*$的OLS估計是第8小題的相減結果，再減去「2014年台北市減去新北市的老人平均活動力」，其中後者反應了兩市立足點的差異（即$\alpha_1$）。

<div class="answer">
Consider the comparison between Taipei and New Taipei city in 2014, which is the comparsion across the following two groups: $D1=1,B1=0,X=0$ and $D1=0, B1=0, X=0$. The comparison is
$$\begin{array}{lcl}
E(y|D1=1,B1=0,X=0)-E(y|D1=0, B1=0, X=0)&=&(\beta_0+\alpha_1)\\
& & -(\beta_0)\\
&=& \alpha_1
\end{array}$$
Deduct it from the previous question's comparison gives us $\beta^*$. Our worry of the comparison that might contain the initial difference between these two cities is removed. 
</div>

（10）很明顯的DiD需要假設：在「不存在福利政策下」，「2014年雙北平均差異」及「2016年雙北平均差異」要相同，即立足點差異不因時間改變。若2016年雙北老年人平均年齡差距拉大（假設北市較高），說明DiD估計會有偏誤。要如何決解？

<div class="answer">
The direct way to fix the problem is to limit our comparision further within a certain age group. In other word, put **age** as part of our control variables. 
</div>


## 5 應用
Run the following codes to download **fatality** data frame object
```{r}
load(url("https://github.com/tpemartin/github-data/blob/master/fatality.Rdata?raw=true"))
```

This is the data for the traffic motality study we discussed during the class

To study the effect of Beer Tax Rate (beertax) on Vehicle Fatality Rate (mrall), we consider the following model:$mrall_{it}=beertax_{it}\beta+c_{i}+\delta_{t}+\epsilon_{it}$,where i represents the i-th state from the U.S. and t represents years. Therefore, $c_{i}$ captures the state effect and $\delta_{t}$ captures the time effect.

1. Use pooled regression model for estimating $\beta$ via OLS.

2. Use Fixed effects model for estimating $\beta$ via LSDV.

3. Compared the results from 2 with from 1, why does the sign of $\beta$ estimate flips?

4. Use Random effects model for estimating $\beta$. Is random effect model more appropriate than the pooled regression model? Is random effect model more appropriate than the fixed effect model?